---
title: "Avalon PLOS"
author: "Olya"
output: 
  html_document:
    keep_md: true
---

## Avalon (well-mapped in osm neighbourhood) PLOS
```{r setup, include=FALSE}
library(tidyverse)
library(sf)
```

##### Reading csv files, merging all line-level data in one table based on the common for sidewalk and highway tables columns.
```{r}
sidewalks <- read.csv("/Users/olechka/Documents/R/PLOS_AVALON/sidewalks.csv")
highways <- read_csv("/Users/olechka/Documents/R/PLOS_AVALON/highway.csv")
line_data <- merge(sidewalks, highways, by = c("wkt_geom", "fid", "full_id", "osm_id", "osm_type", "highway", "motor_vehicle", "access", "horse", "foot", "lit", "oneway", "surface", "bicycle", "name"), all = TRUE)
```

##### Now moving to point data and repeatins previous actions - merging all point-level data in one table based on the common for trees, lights, and parking tables columns.
```{r}
trees <- read.csv("/Users/olechka/Documents/R/PLOS_AVALON/trees.csv")
lights <- read.csv("/Users/olechka/Documents/R/PLOS_AVALON/street_lamp.csv")
parking <- read.csv("/Users/olechka/Documents/R/PLOS_AVALON/parking_aisle.csv")
point_data <- merge( trees, lights, by = c("wkt_geom", "fid", "full_id", "osm_id", "osm_type", "lit"), all = TRUE)

point_data_full <- merge( point_data, parking, by = c("wkt_geom", "fid", "full_id", "osm_id", "osm_type"), all = TRUE)
```


##### Performing join of points and lines to create a master df file
```{r}
Avalon_PLOS_df <- full_join(point_data_full, line_data)
head(Avalon_PLOS_df)
```


##### converting df into sf because we weren't able to write shp
```{r}
Avalon_PLOS_sf <- st_as_sf(Avalon_PLOS_df, wkt = "wkt_geom", crs = 4326)
Avalon_PLOS_sf <- Avalon_PLOS_sf %>%   mutate_if(is.character, ~ if_else(. == "NULL", NA_character_, .))
point_data_sf <- st_as_sf(point_data_full, wkt = "wkt_geom", crs = 4326)
line_data_sf <- st_as_sf(line_data, wkt = "wkt_geom", crs = 4326)

plot(st_geometry(point_data_sf), col = 'blue')
plot(st_geometry(line_data_sf), col = 'red', add = TRUE)   
plot(st_geometry(Avalon_PLOS_sf), col = "green")
```

